\section{Introduction}
\label{sec:intro}

\section{Introduction}
\section{Case Study}
\todo[inline]{Explain what the program is meant for}

This document follows a previous work where \texttt{polu} application - which computes the evolution of a pollutant in a given environment - was studied in order to find possibilities of optimization and/or parallelization.

The original \texttt{polu} application works like a heartbeat algorithm with no communication (since it is executed in a single computation node). The algorithm used by the application sees the environment as a discrete mesh (represented by its cells and edges) and loops until the specified time interval is reached. At each iteration of this main loop, the algorithm performs two main steps:

\begin{description}
	\item[Flux Computation] Based on current pollution values for each cell, the flux for each edge is calculated (performed by the \texttt{compute\_flux} function)
	\item[Pollution Update] Using the previously calculated flux values, the pollution for each cell is updated (performed by the \texttt{update} function)
\end{description}

Several changes were performed in the original \texttt{polu} code in order to allow parallelization and/or improve performance. The most important of these optimizations involved changing the data structures (originally implemented as \textit{Arrays-of-Pointers}) to \textit{Arrays-of-Structures} at first, and then to \textit{Structure-of-Arrays}. The former removed the excessive dereferencing caused by deep chains of pointers in the original structures, which effectively reduced data access time and improved locality, and the latter improved the memory access pattern by preventing unnecessary data pieces from being loaded in the core functions.

The goal of this stage is to attempt and an implementation of the same application, using a distributed memory model with MPI, and analyze the scalability of the algorithm when testing with different amounts of processes and nodes.

\Cref{sec:partitioning} describes the main problem in implementing the \texttt{polu} program in distributed memory, and \cref{sec:implementation} proceeds in explaining how it was implemented. Profiling results are presented in \cref{sec:results}, and the final conclusions for this document in \cref{sec:conclusions}. \Cref{sec:summary} summarizes the results achieved so far with this project, which are the motivation for the next and final step.
